abxD01 Paper

### 2015-03-26 PAPER.Rmd
Fix order of x-axis labels; output correlation values

![**Supplemental Figure 1.**](results/figures/supplemental/sfigure1.png)

**Supplemental Figure 1. The inverse of Simpson's diversity index for the community on day 0 is negatively correlated with *C. difficile* colonization on day 1.** Spearman's rho is plotted in the top right corner.

diversity<-read.csv("~/Documents/Github/abxD01/supplemental/topdose_diversity.csv",  header=T)
diversity$nextDayCFU <- log((diversity$nextDayCFU+1), 10)

#correlation of top dose
pval.spear <- c()
cor.spear <- c()
count <- 1
for(i in 4:5){
  cor.spear[count] <- cor.test(diversity[,3],diversity[,i], method="spearman")$estimate
  pval.spear[count] <- cor.test(diversity[,3],diversity[,i], method="spearman")$p.value
  count <- count + 1
}
corr.results <- as.matrix(cbind(cor.spear, pval.spear))
rownames(corr.results) <- names(diversity)[4:5]

# Plot inverse simpson vs. C. difficile levels
color<-rainbow(7)
colors.top <- c(control = "black",
          cipro = color[1],
          clinda = "#FFCC00",
          vanc.625 = "#00CC00",
          strep5 = "#33FFFF",
          cef.5 = color[5],
          amp.5 = "#6600FF",
          metro1 = color[7])

# shapes
pch.top <- c(control = 1,
         cipro = 16,
         clinda = 16,
         vanc.625 = 16,
         strep5 = 16,
         cef.5 = 16,
         amp.5 = 16,
         metro1 = 16)

#these are needed for the legend
legend.labels.top <- c(control = "Control",
            cipro = "Ciprofloxacin",
            clinda = "Clindamycin",
            vanc.625 = "Vancomycin",
            strep5 = "Streptomycin",
            cef.5 = "Cefoperazone",
            amp.5 = "Ampicillin",
            metro1 = "Metronidazole")

par(mar=c(5, 5, 1, 1) +0.1, mgp=c(3, 1, 0))
par(mfrow=c(1,1))

plot(diversity$invsimpson[diversity$expgroup=="control"],
     diversity$nextDayCFU[diversity$expgroup=="control"],
     pch=pch.top[1], col=colors.top[1], xlim=c(0, 25), ylim=c(0,10),  
     cex=1.5, yaxt="n",  xaxt="n", cex.lab=1,
     ylab=as.expression("Log " ~ italic("C. difficile") ~ " CFU/g feces"), xlab="Inverse Simpson Index")

axis(1, las=1, cex.axis=1, at=seq(0,25, by=5))
axis(2, las=1, cex.axis=1, at=seq(0,10, by=1))

for(i in 2:(length(unique(diversity$expgroup)))){
  points(diversity$invsimpson[diversity$expgroup==names(legend.labels.top)[i]],
         diversity$nextDayCFU[diversity$expgroup==names(legend.labels.top)[i]],
         pch=pch.top[i], col=colors.top[i], cex=1.5)
}

legend("topright", inset=0.05, legend=legend.labels.top, pch=pch.top, col=colors.top,  pt.cex=1.5, cex=1, bty="n", y.intersp=1.5)

@@

***Correlation analysis reveals potentially protective bacteria.*** To identify bacterial taxa that could be associated with resistance or susceptibility to *C. difficile*, we measured the correlation between the relative abundance of each species-level OTU on the day of inoculation with the level of *C. difficile* colonization 24 hours later (Figure 2). <needs help> Of the OTUs with an average abundance greater than 0.1%, OTUs with negative a correlation were associated with providing resistance against *C. difficile* (N=XXX40XXX) and those with a positive correlation were associated with susceptibility (N=XXX7XXX) (p<XXX0.05XXX).</needs help> Among various bacterial families, three were consistently negatively correlated with *C. difficile* colonization: *Porphyromonadaceae* (ρ<sub>average</sub>=XXX-0.63XXX), *Lachnospiraceae* (ρ<sub>average</sub>=XXX-0.51XXX), *Ruminococcaceae* (ρ<sub>average</sub>=XXX-0.53XXX). OTUs from the *Proteobacteria* phylum had strong positive relationships with *C. difficile* colonization. These included OTUs associated with the genera *Pseudomonas* (ρ=XXX0.31XXX) and *Escherichia* (ρ=XXX0.56XXX). Other notable OTUs that were found at the extremes of the distribution of correlation scores included those affiliated with the genus *Lactobacillus* (family *Lactobacillaceae*; OTU23, ρ=XXX-0.73XXX), *Alistipes* (family *Rickenellaceae*; OTU20, ρ=XXX-0.78XXX), *Enterorhabdus* (phylum *Actinobacterium*, family *Coriobacteriaceae*; OTU29, ρ=XXX-0.70XXX), and *Streptococcus* (family *Streptococcaceae*; OTU78, ρ=XXX0.36XXX). Overall, these results suggest that several larger groups of related OTUs, particularly within the *Firmicutes* and *Bacteroidetes* phyla, may participate in colonization resistance against *C. difficile*, while members of the *Proteobacteria* phylum may be involved in susceptibility.

>>>>

```{r top_dose_correlation}

@@

***Correlation analysis reveals potentially protective bacteria.*** To identify bacterial taxa that could be associated with resistance or susceptibility to *C. difficile*, we measured the correlation between the relative abundance of each OTU that had an average abundance greater than 0.1% on the day of inoculation with the level of *C. difficile* colonization 24 hours later (Figure 2).

OTUs with negative a correlation were associated with providing resistance against *C. difficile* (N=XXX40XXX) and those with a positive correlation were associated with susceptibility (N=XXX7XXX) (p<XXX0.05XXX).</needs help> Among various bacterial families, three were consistently negatively correlated with *C. difficile* colonization: *Porphyromonadaceae* (ρ<sub>average</sub>=XXX-0.63XXX), *Lachnospiraceae* (ρ<sub>average</sub>=XXX-0.51XXX), *Ruminococcaceae* (ρ<sub>average</sub>=XXX-0.53XXX). OTUs from the *Proteobacteria* phylum had strong positive relationships with *C. difficile* colonization. These included OTUs associated with the genera *Pseudomonas* (ρ=XXX0.31XXX) and *Escherichia* (ρ=XXX0.56XXX). Other notable OTUs that were found at the extremes of the distribution of correlation scores included those affiliated with the genus *Lactobacillus* (family *Lactobacillaceae*; OTU23, ρ=XXX-0.73XXX), *Alistipes* (family *Rickenellaceae*; OTU20, ρ=XXX-0.78XXX), *Enterorhabdus* (phylum *Actinobacterium*, family *Coriobacteriaceae*; OTU29, ρ=XXX-0.70XXX), and *Streptococcus* (family *Streptococcaceae*; OTU78, ρ=XXX0.36XXX). Overall, these results suggest that several larger groups of related OTUs, particularly within the *Firmicutes* and *Bacteroidetes* phyla, may participate in colonization resistance against *C. difficile*, while members of the *Proteobacteria* phylum may be involved in susceptibility.
